Boycotts require significant deterioration in a company's public image to have an impact on their stock prices.

### 2024-04-26 paper/paper.qmd
updates to paper and saving files as parquet

# United Parks and Recreation
cleaned_data_PRKS <- read.csv("../data/analysis_data/cleaned_data_PRKS.csv")
cleaned_data_PRKS <- 
  cleaned_data_PRKS |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_PRKS$Date <- 
  as.Date(cleaned_data_PRKS$Date, format='%Y-%m-%d')

@@

The objective of this paper is to add to existing literature by determining the impact of boycotting on stock prices and the factors involved in its effectiveness, particularly media coverage. The estimand is the change in stock price during the boycott period. The beginning of the boycott is determined by an event which sparks the most media coverage or is mentioned the most in the media. The end of the boycott is determined by mention in the media of a “turning point” for the company where the reason for the boycott is addressed. Since Starbucks is an ongoing boycott at the time of this research, the “end date” is the date research began. The stock prices when the market closed were arbitrarily selected instead of the stock price at open. To investigate, Nike, United Parks and Recreation, and Starbucks are analyzed as large U.S. companies that were famously boycotted during the desired time period. The analysis includes a detailed view of the difference in stock prices overall, during periods of heavy media coverage, and in comparison with market trends. From analysis, the stock prices during the boycott of Nike were found to not be impacted, United Parks and Recreation was heavily impacted, and Starbucks were mildly impacted. The range in impact can be attributed to overall company image, product type, earnings calls, unionization, and social media. This result supports the importance of media coverage, but also the lack of financial repercussions to many boycotted companies.

@@

## United Parks and Recreation

@@

dataALL <- data.frame(
  "Companies" = c("Nike (NKE)", "United Parks and Recreation (PRKS)", "Starbucks (SBUX)"),
  "Days Observed" = c(N_NKE, N_PRKS, N_SBUX),
  "Date Start" = c(Day1_NKE, Day1_PRKS, Day1_SBUX),
  "Date End" = c(Day2_NKE, Day2_PRKS, Day2_SBUX)
)

@@

## Change in Stock Price
The change in stock price over the length of the boycott is observed in @fig-linegraphNKE, @fig-linegraphPRKS, and @fig-linegraphSBUX. These three graphs show a week before the boycott started to the end of the boycott. @fig-linegraphNKE shows how Nike’s stock price was relatively unstable in the first three years of the boycott, with stock prices rising later in 1991 before dropping slightly and then peaking early in 1993 before descending to a low a year later. A progressive cimb from that point to February of 1997 sees a growth of around 750%. @fig-linegraphPRKS shows that United Parks and Recreation had a steep drop in stock price when ‘Blackfish’ released and had a year of poor stability after before dropping drastically. Afterwards the stock price continues to remain unstable.  @fig-linegraphSBUX graphs the stock price of Starbucks and depicts a steep climb in price in November of 2023 with a peak in early December before a steady climb down over the next 5 months.

@@

```{r}
#| label: fig-linegraphPRKS
#| fig-cap: Changes in United Parks and Recreation Stock Price from One Week Before Blackfish was Released to One Week After Their Pledge to End Orca Breeding
#| echo: false

@@

@tbl-starttoendNKE shows Nike had a small decrease in stock price from a week before the beginning of the boycott. There is also an increase in stock price from the day the boycott started to the day it ended. Relative to the stock price, the change is very large as it increased by 335.82%. @tbl-starttoendPRKS shows that there was a significant decrease of 55.3% in United Parks and Recreation's stock price from the beginning to end of the boycott. Starbucks was relatively unchanged from the start to the end of the boycott with a decrease of only 2.5% as shown in @tbl-starttoendSBUX.

@@

```{r}
#| label: tbl-starttoendPRKS
#| tbl-cap: Changes in United Parks & Recreation Stock Price from One Week Before Blackfish was Released, to the Date it was Released, to their Announcement that they will end their Orca Breeding Program
#| echo: false

@@

## Detailed Observation of Change in Stock Price
In @tbl-twoweeksNKEprotest, the daily difference in Nike's stock price folowing the international Global Exchange protest. The largest change is a week after the protest on the 27th with a drop in 5.79% or 35 cents. 
@tbl-twoweeksSBUX shows the daily difference in stock price for two weeks from the announcement of the lawsuit. The stock prices were relatively unaffected immediately following the announcement and November 2nd is responsible for the large spike in price shown in @fig-linegraphSBUX

```{r}
#| label: fig-linegraphNKEfocused
#| fig-cap: Changes in Nike Stock Price from CAIR Boycott to Two Weeks After the Global Exchange Protest in 1997
#| echo: false

cleaned_data_NKE |>
  filter(between(Date, as.Date('1997-03-02'), as.Date('1997-11-01'))) |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "July 12th, 1991 to May 19th, 1998", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "1 month",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(5,9) + 
  theme_minimal()

@@

@fig-spxPRKS compares the stock price of United Parks & Recreation with the S&P 500 Index (SPX). The comparison cannot indicate whether United Parks was doing better than the market based on price, but by comparing the change in price it is clear that overall the market and United Parks were on a fall entering into 2015. However, the steep price drop at United Parks cannot be explained by the market. The gradual decline at the beginning of the boycott also does not resemble the upward trajectory in the market in 2014.

@@

```{r}
#| label: fig-spxPRKS
#| fig-cap: Changes in United Parks and Recreation Stock Price in Comparison with the S&P 500 index
#| echo: false

@@

## Nike: Seemingly unaffected
Nike is the only company with an increase in stock price over the boycott period and it is quite significant at 334%. Despite this overall growth, Nike does not remain unaffected by the boycotts. In 1997 the company suffered a significant decrease in comparison to the overall growth in the market as seen in @graph-djiNKE.

>>>>

# United Parks and Resorts
cleaned_data_PRKS <- read.csv("../data/analysis_data/cleaned_data_PRKS.csv")
cleaned_data_PRKS <- 
  cleaned_data_PRKS |>
  mutate(Date = lubridate::ymd(Date))
cleaned_data_PRKS$Date <- 
  as.Date(cleaned_data_PRKS$Date, format='%Y-%m-%d')

@@

The objective of this paper is to add to existing literature by determining the impact of boycotting on stock prices and the factors involved in its effectiveness, particularly media coverage. The estimand is the change in stock price during the boycott period. The beginning of the boycott is determined by an event which sparks the most media coverage or is mentioned the most in the media. The end of the boycott is determined by mention in the media of a “turning point” for the company where the reason for the boycott is addressed. Since Starbucks is an ongoing boycott at the time of this research, the “end date” is the date research began. The stock prices when the market closed were arbitrarily selected instead of the stock price at open. To investigate, Nike, United Parks and Resorts, and Starbucks are analyzed as large U.S. companies that were famously boycotted during the desired time period. The analysis includes a detailed view of the difference in stock prices overall, during periods of heavy media coverage, and in comparison with market trends. From analysis, the stock prices during the boycott of Nike were found to not be impacted, United Parks and Resorts was heavily impacted, and Starbucks were mildly impacted. The range in impact can be attributed to overall company image, product type, earnings calls, unionization, and social media. This result supports the importance of media coverage, but also the lack of financial repercussions to many boycotted companies.

@@

## United Parks and Resorts

@@

dataALL <- data.frame(
  "Companies" = c("Nike (NKE)", "United Parks and Resorts (PRKS)", "Starbucks (SBUX)"),
  "Days Observed" = c(N_NKE, N_PRKS, N_SBUX),
  "Date Start" = c(Day1_NKE, Day1_PRKS, Day1_SBUX),
  "Date End" = c(Day2_NKE, Day2_PRKS, Day2_SBUX)
)

@@

## Change in Stock Price
The change in stock price over the length of the boycott is observed in @fig-linegraphNKE, @fig-linegraphPRKS, and @fig-linegraphSBUX. These three graphs show a week before the boycott started to the end of the boycott. @fig-linegraphNKE shows how Nike’s stock price was relatively unstable in the first three years of the boycott, with stock prices rising later in 1991 before dropping slightly and then peaking early in 1993 before descending to a low a year later. A progressive cimb from that point to February of 1997 sees a growth of around 750%. @fig-linegraphPRKS shows that United Parks and Resorts had a steep drop in stock price when ‘Blackfish’ released and had a year of poor stability after before dropping drastically. Afterwards the stock price continues to remain unstable.  @fig-linegraphSBUX graphs the stock price of Starbucks and depicts a steep climb in price in November of 2023 with a peak in early December before a steady climb down over the next 5 months.

@@

```{r}
#| label: fig-linegraphPRKS
#| fig-cap: Changes in United Parks and Resorts Stock Price from One Week Before Blackfish was Released to One Week After Their Pledge to End Orca Breeding
#| echo: false

@@

@tbl-starttoendNKE shows Nike had a small decrease in stock price from a week before the beginning of the boycott. There is also an increase in stock price from the day the boycott started to the day it ended. Relative to the stock price, the change is very large as it increased by 335.82%. @tbl-starttoendPRKS shows that there was a significant decrease of 55.3% in United Parks and Resorts's stock price from the beginning to end of the boycott. Starbucks was relatively unchanged from the start to the end of the boycott with a decrease of only 2.5% as shown in @tbl-starttoendSBUX.

@@

```{r}
#| label: tbl-starttoendPRKS
#| tbl-cap: Changes in United Parks & Resorts Stock Price from One Week Before Blackfish was Released, to the Date it was Released, to their Announcement that they will end their Orca Breeding Program
#| echo: false

@@

## Detailed Observation of Change in Stock Price
In @tbl-twoweeksNKEprotest, the daily difference in Nike's stock price following the international Global Exchange protest. The largest change is a week after the protest on the 27th with a drop in 5.79% or 35 cents. For a better understanding of the overall decline in 1997, @fig-linegraphNKEfocused shows changes in Nike's stock price. The company sees growth early in the year until a rapid drop from mid-February to mid-April. Another significant drop is from August until the end of the year.

@tbl-twoweeksSBUX shows the daily difference in stock price for two weeks from the announcement of the lawsuit. The stock prices were relatively unaffected immediately following the announcement and November 2nd is responsible for the large spike in price shown in @fig-linegraphSBUX

```{r}
#| label: fig-linegraphNKEfocused
#| fig-cap: Changes in Nike Stock Price in 1997
#| echo: false

cleaned_data_NKE |>
  filter(between(Date, as.Date('1997-01-01'), as.Date('1998-01-01'))) |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "January 1st, 1997 to December 31st, 1997", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "1 month",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(5,10) + 
  theme_minimal()

```{r}
#| label: fig-linegraphPRKSfocused
#| fig-cap: Changes in United Parks and Resorts During the Month SeaWorld First Announced they Were Impacted by the Release of 'Blackfish'
#| echo: false

cleaned_data_PRKS |>
  filter(between(Date, as.Date('2014-08-01'), as.Date('2014-09-01'))) |>
  ggplot(mapping = aes(x = Date, y = Close)) +
  geom_line(color="#0000c2") +
  labs(x = "August 1st, 2014 to September 1st, 2014", y = "Stock Price at Closing",) +
  scale_x_date(
    breaks = "1 week",
    labels = scales::label_date_short(
      format = c("%y", "%b"),
      sep = "\n")) +
  ylim(17,30) + 
  theme_minimal()

@@

@fig-spxPRKS compares the stock price of United Parks & Resorts with the S&P 500 Index (SPX). The comparison cannot indicate whether United Parks was doing better than the market based on price, but by comparing the change in price it is clear that overall the market and United Parks were on a fall entering into 2015. However, the steep price drop at United Parks cannot be explained by the market. The gradual decline at the beginning of the boycott also does not resemble the upward trajectory in the market in 2014.

@@

```{r}
#| label: fig-spxPRKS
#| fig-cap: Changes in United Parks and Resorts Stock Price in Comparison with the S&P 500 index
#| echo: false

@@

## Nike: Seemingly unaffected
Nike is the only company with an increase in stock price over the boycott period and it is quite significant at 334%. Despite this overall growth, Nike does not remain unaffected by the boycotts. In 1997 the company suffered a significant decrease in comparison to the overall growth in the market as seen in @graph-djiNKE. The steep drop in February
